PEACE: {\underline P}arallel {\underline E}nvironment for {\underline A}ssembly and {\underline C}lustering of Gene {\underline E}xpression

### 2010-04-18 peace/papers/NAR_web_server_2009/paper.tex
Upated to paper

\begin{figure}
  \centerline{\includegraphics[width=3in]{workflow.eps}}
  \caption{Overview of the procedure for clustering and analysis using
    PEACE.}\label{fig:workflow}
\end{figure}

@@

\peace\/ has been tested on both simulated and real data from NGS and
Sanger Sequencing technologies, comparing results against those
produced by the \wcd\/ clustering tool \wc{(version 0.5.1)}\cite{Hazelhurst08a} and the
\capthree\/ assembly tool \cite{Huang99} (the latter of which
implicitly calculates a clustering in the process of assembly).  For our
simulation tests we used the {\bf ESTSim} tool \cite{Hazelhurst03} to
generate simulated transcript fragments of varying length under
different models of error (Supplementary Materials, Section~C.1),
generating the fragments from the list of 100 zebra fish genes used in
the \wcd\/ testing \cite{Hazelhurst08a}.  Tool parameters were taken
to match, as closely as possible, those used in the \wcd\/ study (see
Supplementary Materials).  Our primary methods of quality assessment
were {\it sensitivity} (the fraction of fragment pairs from the same
gene that were correctly clustered together) and {\it Type 1 error}
(the fraction of genes that were divided between clusters)
\cite{Wang04,Hazelhurst08a}.  Specificity was not an issue in
simulated tests, as both \peace\/ and \wcd\/ were completely successful
in separating unrelated ESTs (save when dealing with recently
duplicated genes).  Other measurements are also discussed in the
supplementary materials.  In Figure~\ref{SeT1} (left), reflecting
tests done on simulated Sanger EST sets, we see a
significant improvement of \peace\/ (blue) over \wcd\/ (green) and
\capthree\/ (black) in its sensitivity to sequencing errors, while in
Figure~\ref{SeT1} (right) we see a comparable improvement in Type 1
error.  In Figure~\ref{singletons} we look at the number of singleton
clusters (fragments not joined to any cluster), which should not occur
in our simulated sets.  We again see significant improvement in \peace.
The sequential runtime of \peace\/ is slightly improved over that of
\wcd\/ (see Supplementary Materials, Figure~S4).

\begin{figure}
  \centerline{\includegraphics[width=3.35in]{SeT1.eps}}
  \caption{Comparisons of Sensitivity and Type 1 error, based on the
    average over 30 simulated Sanger Sequence ESTs sets derived from 100 zebra
    fish genes  (see
    Supplementary Materials, Section C, for more details).  Blue/Solid
    = \peace, Green/Dash = \wcd, Black/Dot-Dash = \capthree; vertical
    tics = 95\% confidence intervals on estimates.  Intervals are not
    presented for Type 1 error due to the extremely small
    variance.}\label{SeT1}
\end{figure}

\begin{figure}
  \centerline{\includegraphics[width=3in]{singletons.eps}}
  \caption{Average number of fragments flagged as singletons by each tool
    when run on the simulated sequences; correct answer in all cases is
    zero.  Blue/Solid = \peace, Green/Dash = \wcd,
    Black/Dot-Dash = \capthree.}\label{singletons}
\end{figure}

>>>>

\begin{figure}
  \centerline{\includegraphics[width=3in]{pics.d/workflow.eps}}
  \caption{Overview of the procedure for clustering and analysis using
    PEACE.}\label{fig:workflow}
\end{figure}

@@

\peace\/ has been tested on both simulated and real data from NGS and
Sanger Sequencing technologies, comparing results against those
produced by the \wcd\/ clustering tool \mc{(version
  0.5.1)}\cite{Hazelhurst08a} and the \capthree\/ assembly tool
\cite{Huang99} (the latter of which implicitly calculates a clustering
in the process of assembly).  For our simulation tests we used the
{\bf ESTSim} tool \cite{Hazelhurst03} to generate simulated transcript
fragments of varying length under different models of error
(Supplementary Materials, Section~C.1), generating the fragments from
the list of 100 zebra fish genes used in the \wcd\/ testing
\cite{Hazelhurst08a}.  Tool parameters were taken to match, as closely
as possible, those used in the \wcd\/ study (see Supplementary
Materials).  \mc{The most important method of quality assessment is
  {\it sensitivity} (the fraction of fragment pairs from the same gene
  that were correctly clustered together).  But we also look at the
  {\it Jaccard Index} (which balances sensitivity with the number of
  false positives by looking at the ratio of true positives to the sum
  of true positives, false positives, and true negitives), {\it Type 1
    error} (the fraction of genes that were divided between clusters),
  and {\it Type 2 error} (the fraction of clusters containing two or
  more genes).  Specificity was not an issue in simulated tests, as
  both \peace\/ and \wcd\/ were completely successful in separating
  unrelated ESTs (save when dealing with recently duplicated genes).
  Other measurements are also discussed in the supplementary
  materials.  In Figure~\ref{SeJiT1T2} (left) we see these four tests,...}

% , reflecting tests done on
%   simulated Sanger EST sets, we see a significant improvement of
%   \peace\/ (blue) over \wcd\/ (green) and \capthree\/ (black) in its
%   sensitivity to sequencing errors, while in Figure~\ref{SeT1} (right)
%   we see a comparable improvement in Type 1 error.  In
%   Figure~\ref{singletons} we look at the number of singleton clusters
%   (fragments not joined to any cluster), which should not occur in our
%   simulated sets.  We again see significant improvement in \peace.
%   The sequential runtime of \peace\/ is slightly improved over that of
%   \wcd\/ (see Supplementary Materials, Figure~S4).

\begin{figure}
  \centerline{\includegraphics[width=3.35in]{pics.d/SeJiT1T2.40.eps}}
  \caption{Comparisons of Sensitivity and Type 1 error, based on the
    average over 30 simulated Sanger Sequence ESTs sets derived from 100 zebra
    fish genes  (see
    Supplementary Materials, Section C, for more details).  Blue/Solid
    = \peace, Green/Dash = \wcd, Black/Dot-Dash = \capthree; vertical
    tics = 95\% confidence intervals on estimates.  Intervals are not
    presented for Type 1 error due to the extremely small
    variance.}\label{SeJiT1T2}
\end{figure}
